---
title: "Week 7 report Group 8"
format: html
---

# PCA tidyverse style on gravier data

# Tasks

| Task                | Responsible | Done? |
|---------------------|-------------|-------|
| Create this file    | Dana        | Yes   |
| Load libraries      | Vakare      | Yes   |
| Load data           | Dana        | Yes   |
| Clean data          | Greta       | No    |
| Produce first plot  | Dana        | No    |
| Produce middle plot | Vakare      | No    |
| Reproduce last plot | Greta       | No    |

# Load libraries

```{r}
library("tidyverse")
library("here")
library("gitcreds")
library("usethis")
```

# Download data the first time and save it ONLY RUN THIS ONCE IF YOU DO NOT HAVE THE DATA

```{r}
# target_url <- "https://github.com/ramhiser/datamicroarray/raw/master/data/gravier.RData"
# output_file <- "data/_raw/gravier.RData"
# curl::curl_download(url = target_url, destfile = output_file)
```

# Load data

```{r}
load(file = here("data/_raw/gravier.RData")) # loads as gravier
```

# Clean data

```{r}
```

# First plot

```{r}
```

# Second plot

```{r}
```

# Third plot

```{r}
```
